---
title: "R Notebook"
author: eyan
date: 23/01/2021
output: html_notebook
---
```{r}
library('tidyverse')
library('data.table')
```
```{r}
url <- 'http://bit.ly/CarreFourDataset'
sup_market <- read.csv(url)
head(sup_market)
```
```{r}
tail(sup_market)
```
```{r}
colSums(is.na(sup_market))
```
No null values were found
```{r}
dim(sup_market)
sup_market[duplicated(sup_market)]

```
No duplicates
```{r}
summary(sup_market)
```
```{r}
invoice         <-sup_market$Invoice.ID
branch          <-sup_market$Branch
customer.type   <-sup_market$Customer.type
gender          <-sup_market$Gender
product.line    <-sup_market$Product.line
unit.price      <-sup_market$Unit.price
quantity        <-sup_market$Quantity
tax             <-sup_market$Tax
date            <-sup_market$Date
time            <-sup_market$Time
payment         <-sup_market$Payment
cogs            <-sup_market$cogs
gross.margin    <-sup_market$gross.margin.percentage
gross.income    <-sup_market$gross.income
rating          <-sup_market$Rating
total           <-sup_market$Total
```
```{r}
out <- boxplot(unit.price)$out
print(paste(length(out),'outliers'))
```
```{r}
out <- boxplot(quantity)$out
print(paste(length(out),'outliers'))
```
```{r}
out <- boxplot(tax)$out
print(paste(length(out),'outliers'))
```
```{r}
out <- boxplot(cogs)$out
print(paste(length(out),'outliers'))
```
```{r}
out <- boxplot(gross.income)$out
print(paste(length(out),'outliers'))
```
```{r}
out <- boxplot(total)$out
print(paste(length(out),'outliers'))
```
The number of outliers was either 9 or 0. It can be assumed the outliers in the dataset occured from the same customers.The outliers will stay so as not to form a biased analysis on customers with specific spending habits.

```{r}
load('kira.RData')
```
```{r}
univar('Branch',sup_market,'category')
```
Most customers bought from branch A, while the fewest were from branch C
```{r}
univar('Customer.type',sup_market,'category')
```
The type of customers in the dataset were the same number
```{r}
univar('Gender',sup_market,'category')
```
The genders were balanced in a 1:1 ratio
```{r}
univar('Product.line',sup_market,'category')
```
Most products bought were fashion accessories then food and beverages
```{r}
univar('Quantity',sup_market,'category')
```
Most products had a rating of 10
```{r}
univar('Payment',sup_market,'category')
```
Most customers paid via E-wallet while the least paid via credit card
```{r}
univar('Unit.price',sup_market)
```
Unit.price ranged from 99.96 to 10.08, and most prices lie between 15.279 and 97.222
The average price was at 55.67, most items had a price of 83.77.

```{r}
univar('Tax',sup_market)
```
Taxes ranged from 49.65 to 0.5085, and most taxes lie between 1.95 and 39.16
The average tax was at 15.379, most customers had a tax of 39.48.
the graph is mostly skewed since most customers lie on the left side.
```{r}
univar('cogs',sup_market)
```

```{r}
univar('gross.income',sup_market)
```
gross income ranged from 49.65 to 0.5085, and most gross incomes lie between 1.95 and 39.16
The average gross income was at 15.379, most customers had a gross income of 39.48.
the graph is mostly skewed since most customers are average with a few on the high end
Taxes and gross income are directly proportional
```{r}
univar('Total',sup_market)
```
Totals ranged from 10.67 and 1042.65 , and most Totals lie between 41.07 and 822.49
The average Total was at 15.379, most customers had a Totals of 829.08.
the graph is mostly skewed since most customers are average spenders with a few on the high end

# Reducing the number of features using PCA
```{r pca}
# importing the necessary library for PCA
library('FactoMineR')
library('ggbiplot')
```
```{r}
label_encode <- function (column,dataframe){
  data <- dataframe[[column]]
  new_data <- as.integer(factor(data))
  # print(new_data)
  return (new_data)
}
supermarket.org <- sup_market[,-c(1,9,10)]

```
```{r}
# Label encoding columns
names_ <- c('Branch','Customer.type','Gender','Product.line','Payment')
for(nm in names_){
  supermarket.org[nm] <- label_encode(nm,supermarket.org)
}
supermarket <- supermarket.org
head(supermarket)
```
```{r}
# Geting the principle components from the supermarkets dataset.
results <- prcomp(supermarket[,-c(10)], center = TRUE, scale. = TRUE)
summary(results)
```
The first principal component has the largest proportion explaining 41% of the variance in the data.
The first 9 components explain the all variance in the data, making the last 3 obsolete
```{r}
screeplot(results, type = "l", npcs = 13, main = "Screeplot of the first 12 Principal components")
abline(h = 0.9, col="red", lty=5)
```
The first 9 principal components explain the most in the data, explainig 99% of the variance. These are the most important features in the dataset.

```{r}
ggbiplot(results,labels.size = 5, alpha = 0.5)
```
Unit price,gross income,tax and quantity contribute to PC1
Payment,gender,Product line and customer type contribute to PC2
```{r}
clusters <- function (name,pc1=1,pc2=2){
  cl_list <- factor(sup_market[[name]])
  # species <- cl_list$species
  ggbiplot(results,,choices=c(pc1,pc2),ellipse = TRUE,groups = cl_list, obs.scale = 1, var.scale = 1,labels.size = 5, alpha = 0.5)
}
clusters('Gender')
```
The clusters of the gender show:
  Males are clustered at the top characterised by all features but the branch, while females are clustered at the bottom with unitprice,tax,gross income,rating and branch influencing their buying.
```{r}
clusters('Gender',3,4)
```
Since the 3rd and 4th princpal components explain lower variances than the first two, i doesnt do that good a job at showing the pattern shown in the previous plot. although it does give a hint on where the different genders lie.
```{r}
clusters('Gender',11,12)
```
The last 2 principle components perform even worse since they ecplain so little of the variance in the dataset.


# Reducing the attributes using Feature Selection
## Method 1 filter method: high correlation filter
```{r}
library('caret')
library('corrplot')
```
```{r}
supermarket_ <- supermarket.org[,-c(10)]
corr_ <- cor(supermarket_)
corr_
```
High correlations from tax,gross income ,cogs and total
```{r}
high_cor <- findCorrelation(corr_, cutoff=0.75)
high_cor
```

```{r}
# Removing the highly correlated features in index (9,12,7)
final <- supermarket_[-high_cor]
```
```{r}
# Plotting the correlations for comparison
par(mfrow = c(1,2))
corrplot(corr_, order = "hclust")
corrplot(cor(final), order = "hclust")
```
Feature selection removes the most correlated features from the dataset in this case were left with 9 features that are not overly correlated

## Method 2 : Wrapper Methods
```{r}
library('clustvarsel')
library('mclust')
```
```{r}
# Sequential forward greedy search using the default number of groups
out <- clustvarsel(supermarket_)
out
```
```{r}
final2 <- supermarket_[,out$subset]
head(final2)
```
Using clustvarsel 3 groups were taken as important in creating the clusters of the dataset
```{r}
mod <- Mclust(final2)
summary(mod)
```
There were achieved 7 clusters in the selected subset of data
```{r}
plot(mod,c("classification"))
```
The unsupervised clustering shows 7 clusters from the 3 attributes selected